Lost in Translation: \\ The function--sheaf dictionary for smooth commutative group schemes over finite fields

### 2014-05-16 dictionary.tex
Adding material on spectral sequences

\date{\today}
\author{Clifton Cunningham}
\email{cunning@math.ucalgary.ca}
\address{University of Calgary}
\author{David Roe}
\email{roed.math@gmail.com}
\address{Pacific Institute for the Mathematical Sciences, University of Calgary}
\subjclass[2010]{14F05 (primary), 14L15 (secondary)}
\keywords{sheaf--function dictionary, geometrization, quasicharacter sheaves, commutative group schemes}

%%%%%%%%%%%%%%%%% PACKAGES %%%%%%%%%%%%%%%%%
% Encodings
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% AMS packages
\usepackage{amsmath, amsthm, amssymb, todonotes, showkeys}
\usepackage[alphabetic]{amsrefs}
% Fonts
\usepackage{mathrsfs, yfonts}
\usepackage{geometry}
% TikZ
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,calc,matrix}
\usepackage{tikz-cd}
% Hyperrefs
\usepackage{hyperref}

@@

@@

\begin{abstract}
  We extend Deligne's function--sheaf dictionary from
  connected commutative algebraic groups over finite fields to smooth
  commutative group schemes over finite fields.
  To do so, we introduce the monoidal category of quasicharacter sheaves on
  smooth commutative group schemes $G$ over $\mathbb{F}_q$
  and show that the Trace of Frobenius is a functorial surjection
  from the group of isomorphism classes of quasicharacter sheaves on $G$
  onto the character group of $G(\mathbb{F}_q)$.
  If $G$ is connected, the kernel of this surjection is trivial.
  When $G$ is etale, however, the kernel is generally non-trivial,
  leading to the interesting notion of invisible quasicharacter sheaves.
  We find a classification of invisible quasicharacter sheaves
  on etale commutative group schemes over $\mathbb{F}_q$
  in terms of the cohomology of $G(\mathbb{F}_q)$.
  We also explain how to make invisible quasicharacter sheaves disappear,
  should the notion prove too troubling.
   \end{abstract}

@@

Three examples will illustrate some of the issues that arise when extending the function--sheaf dictionary to etale, commutative group schemes over $\Fq$.
As we show in the paper, $\ell$-adic local systems on $G$ can only be used to geometrize characters of finite order; quasicharacter sheaves, however, provide the geometrization of all characters of $G(\Fq)$.
In particular, we will see how to uniquely geometrize the character $\ZZ \to \EEx$ defined by
$1 \mapsto \ell$, which has unbounded image,
using quasicharacter sheaves on the etale group scheme $\ZZ$ with trivial Galois action.
All quasicharacter sheaves on this group are visible.
Likewise, the dictionary is perfect for the
etale group scheme $\ZZ$ with Galois action determined by $\Frob{}(n) = -n$.
By contract, the etale group scheme $\ZZ \times \ZZ$ with trivial Galois action
 does admit invisible quasicharacter sheaves;
 in particular, there are non-trivial quasicharacter sheaves on this group
with trivial trace of Frobenius.\footnote{Move this example: Define $g: \ZZ\times \ZZ \to \EEx$ by $g((n_1,m_1), (n_2, m_2)) = \ell^{m_1-m_2 + n_1-n_2}$.
Then use the tehniques of Section~\ref{sec:etale} to build a non-trivial character sheaf from the pair $(f=1,g)$ with trivial trace of Frobenius.}

@@

We now begin developing the tools to relate $\QCiso{G}$ to $G(\Fq)^*$.
%
Let $(\gqcs{L},\phi)$ be a Weil sheaf on $G$. Every $g\in G(\Fq)$
determines a point $\bg$ on $\bG$ fixed by $\Frob{G}$ and
therefore an automorphism $\phi_{\bg}$ of $\gqcs{L}_{\bg}$. Set $\trFrob{(\gqcs{L},\phi)}(g) \ceq \trace(\phi_{\bg}, \gqcs{L}_{\bg})$,
and note that if $(\gqcs{L},\phi) \iso (\gqcs{L'},\phi')$ then $\trFrob{(\gqcs{L},\phi)} = \trFrob{(\gqcs{L'},\phi')}$.
 If $\qcs{L} = (\gqcs{L},\mu,\phi)$ is a quasicharacter sheaf
then we will abbreviate $\trFrob{(\gqcs{L},\phi)}$ to $\trFrob{\qcs{L}}$.

\begin{definition}\label{def:trFrob}
The function
\begin{align*}
\TrFrob{G}: \QCiso{G} &\to G(\Fq)^* \\
\qcs{L} &\mapsto t_{\qcs{L}}
\end{align*}
is called \emph{Trace of Frobenius}.\todo{Does this really warrant a declared definition?
Instead, we can refer to this very short section, when needed.}
\end{definition}

@@

\begin{proposition}\label{prop:functorialG}
Trace of Frobenius $\TrFrob{G} : \QCiso{G} \to G(\Fq)^*$ is a natural transformation
between the two additive functors
\begin{align*}
F_1 : G &\mapsto \QCiso{G} \\
F_2 : G &\mapsto G(\Fq)^*
\end{align*}
from the category of smooth commutative group schemes over $\Fq$ to the category of abelian groups.
\end{proposition}

@@

\subsection{Invisible quasicharacter sheaves}\label{sec:invisible}

We will see that the Trace of Frobenius may not provide complete
information about isomorphism classes of quasicharacter sheaves on $G$ when $G$ is not connected.
Our main tool for understanding this phenomenon is the group homomorphism
\begin{equation}\label{2cocycle}
S_{G}: \QCiso{G} \to H^0(\Weil{},H^2(G(\bFq),\EEx))
\end{equation}
defined here.
Let $\qcs{L} = (\gqcs{L},\mu,\phi)$ be a quasicharacter sheaf on $G$.
For each geometric point $x\in G(\bFq)$, choose a basis $\{ v_x \}$ for $\gqcs{L}_x$.
This choice determines a function $g : G(\bFq)\times G(\bFq) \to \EEx$ defined by
\[
\mu_{x,y}(v_{x+y}) = g(x,y) v_x \otimes v_y.
\]
Chasing through the diagram in Condition~\ref{QC.2} we see that
\[
g(x+y,z) g(x,y) = g(x,y+z) g(y,z),
\]
for all $x,y,z\in G(\bFq)$, so $g \in Z^2(G(\bFq),\EEx)$.
It is easy to see that the class of $g$ is independent of the choice made above;
let $s_\mathcal{L} \in H^2(G(\bFq),\EEx)$ the class of $g\in Z^2(G(\bFq),\EEx)$.
It is also easy to see that $s_\mathcal{L} = s_\mathcal{L'}$ when $\qcs{L} \iso \qcs{L'}$, and that $s_{\qcs{L}\otimes \qcs{L'}} = s_{\qcs{L}} \, s_\qcs{L'}$
for every $\qcs{L}, \qcs{L'} \in \QC(G)$.
We claim that $s_\qcs{L}$ is fixed by the action of $\Weil{}$ on $H^2(G(\bFq),\EEx)$
inherited from the action of $\Weil{}$ on $G(\bFq)$.
To see this, we return to the choice $\{ v_x \in \gqcs{L}_x^\times \tq x\in G(\bFq) \}$
and define $f : G(\bFq) \to \EEx$ by
\[
\phi_x(v_{\Frob{G}(x)}) = f(x) v_x.
\]
Observe:
\begin{equation}\label{ftot}
f\vert_{G(\Fq)} = t_\qcs{L},
\end{equation}
which is independent of the choice of $\{ v_x \in \gfqcs{L}_x \tq x\in G(\bFq)\}$.
Chasing through the diagram in Condition~\ref{QC.3} we see that
\begin{equation}\label{nohom}
g(\Frob{G}(x),\Frob{G}(y)) f(x) f(y)  = g(x,y) f(x+y)
\end{equation}
for all $x,y \in G(\bFq)$; thus, the class of $\,^{\Frob{}}g$ coincides with the class of $g$.
This concludes the definition of
\begin{align*}
S_G: \QCiso{G} &\to \Hh^0(\Weil{},\Hh^2(G(\bFq),\EEx))\\
\qcs{L} &\mapsto s_{\qcs{L}}.
\end{align*}

\begin{definition}\label{def:invisible}
Let $\qcs{L}$ be a quasicharacter sheaf on $G$. We say $\qcs{L}$ is \emph{visible} if $s_\qcs{L} \in H^0(\Weil{},H^2(G(\bFq),\EEx))$ is trivial; otherwise, we say $\qcs{L}$ is \emph{invisible}.
\end{definition}

This terminology will be justified later when\todo{Point somewhere.} we show that a non-trivial quasicharacter sheaf $\qcs{L}$ on $G$ is invisible if and only if $\trFrob{\qcs{L}}$ is trivial.

@@

The definition of finite quasicharacter sheaves will require several steps,
beginning with the localization of a category introduced here.
Let us say that a morphism $f : H\to G$ of smooth commutative group schemes over $\Fq$ is a \emph{discrete isogeny}
if it is finite, surjective and \'etale and the action of $\Gal(\bFq/\Fq)$ on the \'etale group
scheme $\ker f \ceq f^{-1}({\bar e)}$ is trivial.
Consider the category $C(G)$ whose objects consist of pairs $(f,\psi)$, where
\begin{itemize}
\item $f : H\to G$ is a discrete isogeny of smooth commutative group schemes over $\Fq$,
\item $\psi : \ker f\to \Aut(V)$ is a representation on a finite-dimensional $\EE$-vector space $V$.
\end{itemize}
A morphism $(f,\psi) \to (f',\psi')$ in $C(G)$ is a pair $(g,T)$, where
\begin{itemize}
\item $g : H' \to H$ is a morphism of schemes such that $f' = f\circ g$,
\item $T : V\to V'$ is a linear transformation, equivariant for the action of
$\ker f'$ on $V'$ by $\psi'$ and the action of $\ker f'$ on $V$ by $\psi \circ \alpha$.
\[
\begin{tikzcd}
V \arrow{d}{T} & \Aut(V) & \arrow[swap]{l}{\psi} \ker f \arrow[hook]{r} & H \arrow{r}{f} & \arrow[equal]{d} G\\
V' & \Aut(V') & \arrow[swap]{l}{\psi'} \arrow[swap]{u}{g\vert_{\ker f'}} \ker f' \arrow[hook]{r} & \arrow[swap]{u}{g} H' \arrow{r}{f'} & G
\end{tikzcd}
\]
\end{itemize}

@@

\begin{proof}
The key point is that pullback along $g$ takes discrete isogenies to discrete isogenies,
and maps over $G$ to maps over $J$, as pictured below.
\[
\begin{tikzcd}[row sep={{{{40,between origins}}}},column sep={{{{50,between origins}}}}]
%\ V \arrow{dd}{T} &
\Aut(V) && \arrow[swap]{ll}{\psi} \ker f \arrow[hook]{rr} \arrow[leftarrow]{dd} && H \arrow{rr}{f} \arrow[leftarrow]{dd} && \arrow[equal]{dd} G\\
% &
& \arrow{ul} \ker f\times_G J \arrow{ur} \arrow[hook, crossing over]{rr} && H\times_G J \arrow{ur} \arrow[crossing over]{rr} && J \arrow[swap]{ur}{g} & \\
%V' &
 \Aut(V') && \arrow[swap,pos=0.25]{ll}{\psi'} \ker f'  \arrow[hook]{rr}
&& H' \arrow[pos=0.25]{rr}{f'} && G\\
% &
 & \arrow{ul} \ker f'\times_G J \arrow[crossing over]{uu} \arrow{ur} \arrow[hook]{rr}
 && \arrow[crossing over]{uu} H'\times_G J \arrow{ur} \arrow{rr} && \arrow[equals, crossing over]{uu} J \arrow[swap]{ur}{g} &
\end{tikzcd}
\]
In this form we could treat $g^*$ as a functor $C(J)\to C(G)$,
were it not for the fact that pullback is defined only up to isomorphism.
In the localized categories, this ambiguity disappears.
\end{proof}

@@

\begin{corollary}
 If $G$ is a connected, commutative algebraic group over $\Fq$ then
 \[
  \TrFrob{G} : \QCiso{G} \to G(\Fq)^*
 \]
 is an isomorphism of groups.
\end{corollary}
\begin{proof}
 It suffices to observe that the group homomorphism
 $\TrFrob{G} : \QCiso{G} \to G(\Fq)^*$
 is surjective because each character $\chi \in G(\Fq)^*$ determines a finite quasicharacter
 sheaf $(\Lang,\chi^{-1})$ such that Trace of Frobenius of the quasicharacter sheaf
 $B_G (L_G(\Lang,\chi^{-1}))$ is $\chi$.
\end{proof}

As Proposition~\ref{prop:connected} shows,
when $G$ is a connected algebraic group over $\Fq$,
it is appropriate to replace quasicharacter sheaves on $G$
with the conceptually simpler category $\QCf(G)$ of finite quasicharacter sheaves on $G$.
In this context, finite quasicharacter sheaves may also be apprehended as
\emph{certain} $\ell$-adic characters of fundamental group $\pi_1(G,{\bar e})$.
As explained in \cite{kamgarpour:09a}*{\S2},
the relevant characters of $\pi_1(G,{\bar e})$ are precisely those
that factor through a particular quotient of $\pi_1(G,{\bar e})$
denoted there by $\Pi_\text{disc}(G)$.
Using this fact, one can show that the category of discrete isogenies to $G$ is a Galois category,
in the sense of \cite{grothendieck:SGA1}*{Expos\'e V, \S 4}.
Moreover, since $\Pi_\text{disc}(G) \iso G(\Fq)$ \cite{kamgarpour:09a}*{App'x B},
the fundamental functor attached to this Galois category can be used to produce an equivalence
between the category of finite quasicharacter sheaves on $G$ and
the category of one-dimensional representations of $G(\Fq)^*$, thus
providing an alternate proof that $\QCiso{G} \iso G(\Fq)^*$
when $G$ is a connected algebraic group over $\Fq$.

In principle, something similar is possible in the more general
context of this paper, where $G$ is a $\QC$-scheme, to which we now return.
In this generality, however, neither of the full subcategories $\QCf(G) \subset \QCb(G) \subset \QC(G)$ are equivalences;
see the end of Section~\ref{sec:snake}.
Nevertheless, every quasicharacter sheaf $(\gqcs{L},\mu,\phi)$ on $G$
determines a Weil sheaf $(\gqcs{L},\phi)$ on $G$
so if we choose a geometric point ${\bar g}_x$ in each component $\bG^x$ of $G$
then this choice can be used to convert
the Weil sheaf $(\gqcs{L},\phi)$ into an $\ell$-adic character of $\prod_{x\in \pi_0(\bG)}\W(\bG^x, {\bar g}_x)$.
However, it seemed cumbersome to manage various
choices of families of base points and track the action
of $\Gal(\bFq/\Fq)$ on this (generally) infinite product of
Weil groups and also difficult to identify the relevant quotient of this group.
Viewed in this light, Definition~\ref{def:QC} is comparatively simple.

\subsection{Quasicharacter sheaves, revisited}

\'Etale group schemes form a counterpoint to connected groups,
since the component group of a smooth commutative group scheme
is an \'etale group scheme \cite{vdGeer-Moonen:AbelianVarieties}*{III, \S 4}.
However, the function--sheaf dictionary is substantially differentl in the two cases.
To see why, we begin by specializing the definition of quasicharacter sheaves to the case
of etale commutative group schemes,
using the equivalence $G \mapsto G(\bFq)$
from the category of \'etale group schemes over $\Fq$ to the category of groups equipped
with an action of $\Gal(\bFq/\Fq)$, continuous for the discrete topology on the group.
Under this equivalence, a quasicharacter sheaf on an etale commutative group scheme $G$ over $\Fq$ is given by the data of:
\begin{enumerate}
 \labitem{(qc.1)}{qc.1} an indexed set of one-dimensional
  $\EE$-vector spaces $\gqcs{L}_x$, as $x$ runs over
  $G(\bFq)$;

\subsection{Group Cohomology}

\begin{proposition} \label{prop:etale-iso}
	If $G$ is an \'etale commutative group scheme over $\Fq$
  	then
  	\[
  	\QCiso{G} \iso \Hh^2(G(\bFq)\rtimes \Weil{}, \EEx),
  	\]
  	functorially,
  	in the sense of Proposition~\ref{prop:functorialG}.
\end{proposition}

The proof of Proposition~\ref{prop:etale-iso} will require a few lemmas,
which we assemble now.\todo{This section is undergoing massive reconstruction. Don't read.}
%
%We begin with a simple result relating duals, invariants and coinvariants.

>>>>

\date{\today}
\author{Clifton Cunningham}
\email{cunning@math.ucalgary.ca}
\address{University of Calgary}
\author{David Roe}
\email{roed.math@gmail.com}
\address{Pacific Institute for the Mathematical Sciences, University of Calgary}
\subjclass[2010]{14F05 (primary), 14L15 (secondary)}
\keywords{function--sheaf dictionary, geometrization, quasicharacters, quasicharacter sheaves, smooth commutative group schemes}

%%%%%%%%%%%%%%%%% PACKAGES %%%%%%%%%%%%%%%%%
% Encodings
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% AMS packages
\usepackage{amsmath, amsthm, amssymb, todonotes}% showkeys}
\usepackage[alphabetic]{amsrefs}
% Fonts
\usepackage{mathrsfs, yfonts}
\usepackage{geometry}
% TikZ
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,calc,matrix}
\usepackage{tikz-cd}
% Hyperrefs
\usepackage{hyperref}

@@

\newcommand{\lQC}{{\mathcal{Q\hskip-0.8pt C}^2}}
\newcommand{\lQCiso}[1]{\lQC(#1)_{/\text{iso}}}
\newcommand{\lTrFrob}[1]{\TrFrob{#1}^2}

@@

\begin{abstract}
  We extend Deligne's function--sheaf dictionary from
  connected commutative algebraic groups over finite fields to smooth
  commutative group schemes over finite fields.
  To do so, we introduce the monoidal category of quasicharacter sheaves on
  smooth commutative group schemes $G$ over $\mathbb{F}_q$
  and show that the Trace of Frobenius is a functorial surjection
  from the group of isomorphism classes of quasicharacter sheaves on $G$
  onto the character group of $G(\mathbb{F}_q)$.
  If $G$ is connected, the kernel of this surjection is trivial.
  In $G$ is etale, however, the kernel is generally non-trivial,
  leading to the interesting notion of invisible quasicharacter sheaves.
  In this paper we find a cohomological classification of invisible quasicharacter sheaves on all smooth commutative group schemes over $\mathbb{F}_q$,
  and also explain how to make invisible quasicharacter sheaves disappear,
  should the notion prove too troubling.
   \end{abstract}

@@

Three examples will illustrate some of the issues that arise when extending the function--sheaf dictionary to smooth commutative group schemes over $\Fq$.
As we show in the paper, $\ell$-adic local systems on $G$ can only be used to geometrize characters of finite order; quasicharacter sheaves, however, provide the geometrization of all characters of $G(\Fq)$.
In particular, we will see how to uniquely geometrize the character $\ZZ \to \EEx$ defined by
$1 \mapsto \ell$, which has unbounded image,
using quasicharacter sheaves on the etale group scheme $\ZZ$ with trivial Galois action.
All quasicharacter sheaves on this group are visible.
Likewise, the dictionary is perfect for the
etale group scheme $\ZZ$ with Galois action determined by $\Frob{}(n) = -n$.
By contract, the etale group scheme $\ZZ \times \ZZ$ with trivial Galois action
 does admit invisible quasicharacter sheaves -- there are non-trivial quasicharacter sheaves on this group
with trivial trace of Frobenius.\todo{Under construction}
(Move this example: Define $g: \ZZ\times \ZZ \to \EEx$ by $g((n_1,m_1), (n_2, m_2)) = \ell^{m_1-m_2 + n_1-n_2}$.
Then use the tehniques of Section~\ref{sec:etale} to build a non-trivial character sheaf from the pair $(f=1,g)$ with trivial trace of Frobenius.)

@@

We now begin developing the tools to relate $\QCiso{G}$ to $G(\Fq)^*$.
%
Let $(\gqcs{L},\phi)$ be a Weil sheaf on $G$. Every $g\in G(\Fq)$
determines a point $\bg$ on $\bG$ fixed by $\Frob{G}$ and
therefore an automorphism $\phi_{\bg}$ of $\gqcs{L}_{\bg}$. Set
\begin{equation}\label{trWeil}
%\begin{equation}\label{trWeil}
\trFrob{(\gqcs{L},\phi)}(g) \ceq \trace(\phi_{\bg}, \gqcs{L}_{\bg})
%\end{equation}
\end{equation}
and note that if $(\gqcs{L},\phi) \iso (\gqcs{L'},\phi')$ then $\trFrob{(\gqcs{L},\phi)} = \trFrob{(\gqcs{L'},\phi')}$.
 If $\qcs{L} = (\gqcs{L},\mu,\phi)$ is a quasicharacter sheaf
then we will abbreviate $\trFrob{(\gqcs{L},\phi)}$ to $\trFrob{\qcs{L}}$.
%\begin{definition}\label{def:trFrob}
We will refere to the function
\begin{align*}
\TrFrob{G}: \QCiso{G} &\to G(\Fq)^* \\
\qcs{L} &\mapsto t_{\qcs{L}}: g\mapsto \trace(\phi_{\bg}, \gqcs{L}_{\bg})
\end{align*}
as the \emph{Trace of Frobenius} for $G$.
%\todo{Does this really warrant a declared definition?
%Instead, we can refer to this very short section, when needed.}
%\end{definition}

@@

\begin{proposition}\label{prop:functorialG}
Trace of Frobenius $\TrFrob{G} : \QCiso{G} \to G(\Fq)^*$ is a natural transformation
between the two contravariant additive functors
\begin{align*}
F_1 : G &\mapsto \QCiso{G} \\
F_2 : G &\mapsto G(\Fq)^*
\end{align*}
from the category of smooth commutative group schemes over $\Fq$ to the category of abelian groups.
\end{proposition}

@@

@@

The definition of finite quasicharacter sheaves will require several steps,
beginning with the localization of a category introduced here.
Let us say that a morphism $f : H\to G$ of smooth commutative group schemes over $\Fq$ is a \emph{discrete isogeny}
if it is finite, surjective and \'etale and the action of $\Gal(\bFq/\Fq)$ on the \'etale group
scheme $\ker f = f^{-1}({\bar e)}$ is trivial.
Consider the category $C(G)$ whose objects consist of pairs $(f,\psi)$, where
\begin{itemize}
\item $f : H\to G$ is a discrete isogeny of smooth commutative group schemes over $\Fq$,
\item $\psi : \ker f\to \Aut(V)$ is a representation on a finite-dimensional $\EE$-vector space $V$.
\end{itemize}
A morphism $(f,\psi) \to (f',\psi')$ in $C(G)$ is a pair $(g,T)$, where
\begin{itemize}
\item $g : H' \to H$ is a morphism of schemes such that $f' = f\circ g$,
\item $T : V\to V'$ is a linear transformation, equivariant for the action of
$\ker f'$ on $V'$ by $\psi'$ and the action of $\ker f'$ on $V$ by $\psi \circ \alpha$.
\[
\begin{tikzcd}
V \arrow{d}{T} & \Aut(V) & \arrow[swap]{l}{\psi} \ker f \arrow[hook]{r} & H \arrow{r}{f} & \arrow[equal]{d} G\\
V' & \Aut(V') & \arrow[swap]{l}{\psi'} \arrow[swap]{u}{g\vert_{\ker f'}} \ker f' \arrow[hook]{r} & \arrow[swap]{u}{g} H' \arrow{r}{f'} & G
\end{tikzcd}
\]
\end{itemize}

@@

\begin{proof}
The key point is that pullback along $g$ takes discrete isogenies to discrete isogenies,
and maps over $G$ to maps over $J$, as pictured below.
\[
\begin{tikzcd}[row sep={{{{40,between origins}}}},column sep={{{{50,between origins}}}}]
%\ V \arrow{dd}{T} &
\Aut(V) && \arrow[swap]{ll}{\psi} \ker f \arrow[hook]{rr} \arrow[leftarrow]{dd} && H \arrow{rr}{f} \arrow[leftarrow]{dd} && \arrow[equal]{dd} G\\
% &
& \arrow{ul} \ker f\times_G J \arrow{ur} \arrow[hook, crossing over]{rr} && H\times_G J \arrow{ur} \arrow[crossing over]{rr} && J \arrow[swap]{ur}{g} & \\
%V' &
 \Aut(V') && \arrow[swap,pos=0.25]{ll}{\psi'} \ker f'  \arrow[hook]{rr}
&& H' \arrow[pos=0.25]{rr}{f'} && G\\
% &
 & \arrow{ul} \ker f'\times_G J \arrow[crossing over]{uu} \arrow{ur} \arrow[hook]{rr}
 && \arrow[crossing over]{uu} H'\times_G J \arrow{ur} \arrow{rr} && \arrow[equals, crossing over]{uu} J \arrow[swap]{ur}{g} &
\end{tikzcd}
\]
%In this form we could treat $g^*$ as a functor $C(J)\to C(G)$,
%were it not for the fact that pullback is defined only up to isomorphism.
%In the localized categories, this ambiguity disappears.
\end{proof}

@@

\begin{corollary}\label{cor:connected}
 If $G$ is a connected, commutative algebraic group over $\Fq$ then
 \[
  \TrFrob{G} : \QCiso{G} \to G(\Fq)^*
 \]
 is an isomorphism of groups.
\end{corollary}
\begin{proof}
 It suffices\todo{Why is $T_G$ injective?} to observe that the group homomorphism
 $\TrFrob{G} : \QCiso{G} \to G(\Fq)^*$
 is surjective: each character $\chi \in G(\Fq)^*$ determines a finite quasicharacter
 sheaf $(\Lang,\chi^{-1})$ such that Trace of Frobenius of the quasicharacter sheaf
 $B_G (L_G(\Lang,\chi^{-1}))$ is $\chi$.
\end{proof}

As Proposition~\ref{prop:connected} shows,
when $G$ is a connected algebraic group over $\Fq$,
it is appropriate to replace quasicharacter sheaves on $G$
with the conceptually simpler category $\QCf(G)$ of finite quasicharacter sheaves on $G$.
In this context, finite quasicharacter sheaves may also be apprehended as
\emph{certain} $\ell$-adic characters of fundamental group $\pi_1(G,{\bar e})$.
As explained in \cite{kamgarpour:09a}*{\S2},
the relevant characters of $\pi_1(G,{\bar e})$ are precisely those
that factor through a particular quotient of $\pi_1(G,{\bar e})$
denoted there by $\Pi_\text{disc}(G)$.
Using this fact, one can show that the category of discrete isogenies to $G$ is a Galois category,
in the sense of \cite{grothendieck:SGA1}*{Expos\'e V, \S 4}.
Moreover, since $\Pi_\text{disc}(G) \iso G(\Fq)$ \cite{kamgarpour:09a}*{App'x B},
the fundamental functor attached to this Galois category can be used to produce an equivalence
between the category of finite quasicharacter sheaves on $G$ and
the category of one-dimensional representations of $G(\Fq)^*$, thus
providing an alternate proof of Corollary~\ref{cor:connected}.

In principle, something similar is possible in the more general
context of this paper, where $G$ is a smooth commutative group scheme over $\Fq$, to which we now return.
In this generality, however, neither of the full subcategories $\QCf(G) \subset \QCb(G) \subset \QC(G)$ are equivalences;
see the end of Section~\ref{sec:snake}.
Nevertheless, every quasicharacter sheaf $(\gqcs{L},\mu,\phi)$ on $G$
determines a Weil sheaf $(\gqcs{L},\phi)$ on $G$
so if we choose a geometric point ${\bar g}_x$ in each component $\bG^x$ of $G$
then this choice can be used to convert
the Weil sheaf $(\gqcs{L},\phi)$ into an $\ell$-adic character of $\prod_{x\in \pi_0(\bG)}\W(\bG^x, {\bar g}_x)$.
However, it seemed cumbersome to manage various
choices of families of base points and track the action
of $\Gal(\bFq/\Fq)$ on this (generally) infinite product of
Weil groups and also difficult to identify the relevant quotient of this group.
Viewed in this light, Definition~\ref{def:QC} seems comparatively simple.

\section{Spectral sequences}\label{sec:spectral}

%\subsection{Invisible quasicharacter sheaves}\label{sec:invisible}

We will see that the Trace of Frobenius may not provide complete
information about isomorphism classes of quasicharacter sheaves on $G$ when $G$ is not a connected algebraic group.
Our main tool for understanding this phenomenon
is a group homomorphism from $\QCiso{G}$ to $\Hh^2(E^\bullet_G)$, defined in this section. In Section~\ref{sec:restriction} we will see that this homomorphism is an isomorphism for all smooth commutative group schemes over $\Fq$.

[Things I need from this section:
definition of $E^\bullet_G$ (Section~\ref{sec:E});
the short exact sequence attached to $\Hh^2(E^\bullet_G)$ (Section~\ref{sec:SES});
the exactness of all sequences in the diagram below (Section~\ref{sec:front});
the map $S_G$ from $\QCiso{G}$ to $\Hh^2(E^\bullet_G)$ (Section~\ref{sec:S} done) commutes with everything in sight;
that $\Hh^0(\Weil,\Hh^2({\bar G},\EEx))=0$ if $G$ is a connected algebraic group over $\Fq$ (Section~\ref{sec:connected-iso}).]

[Can this section be written in such a way that the effect of removing the fact $\Hh^2(\Weil{}, - ) =0$ can be discerned?]

\subsection{Cohomology}\label{sec:E}

Define $\Hh^2(E^\bullet_G)$.

Do we have
 \[
  \begin{tikzcd}[column sep=30]
    ? \cdots \arrow{r}
%    & \Hh^1(E^\bullet_{A}) \arrow{r}
    & \Hh^2(E^\bullet_{C}) \arrow{r}
    & \Hh^2(E^\bullet_{B}) \arrow{r}
    & \Hh^2(E^\bullet_{A}) \arrow{r}
%    & \Hh^3(E^\bullet_{C}) \arrow{r}
    & \cdots ? \\
  \end{tikzcd}
 \]

[Question: If $C$ is etale, then $\Hh^3(E^\bullet_{C}) =0?$]

\[\Hh^*(E^\bullet_{G}) = \Hh^*({\bar G}\rtimes \Weil{},\EEx)\]

\subsection{A short exact sequence}\label{sec:SES}

 \[
  \begin{tikzcd}
    0 \arrow{r}
    & \Hh^0(\Weil{},\Hh^2({\bar G},\EEx)) \arrow{r}
    & \Hh^2(E^\bullet_G) \arrow{r}
    & \Hh^1(\Weil{},\Hh^1({\bar G},\EEx)) \arrow{r}
    & 0
  \end{tikzcd}
 \]

\subsection{Splitting?}\label{sec:split}

 \[
  \begin{tikzcd}
    0 \arrow{r}
    & \Hh^0(\Weil{},\Hh^2({\bar G},\EEx)) \arrow{r}
    & \Hh^2(E^\bullet_G) \arrow{r}
    \arrow[bend right,dotted]{l}
    & \Hh^1(\Weil{},\Hh^1({\bar G},\EEx)) \arrow{r}
    \arrow[bend right,dotted]{l}
    & 0
  \end{tikzcd}
 \]

\subsection{Front}\label{sec:front}

 \[
  \begin{tikzcd}[row sep=20, column sep=15]
    &  ?  & 0 \arrow{d} & 0\arrow{d} & \\
0 \arrow{r} \arrow{d}   & \Hh^0(\Weil{},\Hh^2(C,\EEx)) \arrow{r} \arrow{d} & \Hh^0(\Weil{},\Hh^2(B,\EEx)) \arrow{r} \arrow{d}
    & \Hh^0(\Weil{},\Hh^2(A,\EEx)) \arrow{d} \arrow{r} & ? \\
    %\arrow{rr}  &   & 0 \\
    ? \arrow{r}
    & \Hh^2(E^\bullet_{C}) \arrow{r} \arrow{d}
    & \Hh^2(E^\bullet_{B}) \arrow{r} \arrow{d}
    & \Hh^2(E^\bullet_{A}) \arrow{r} \arrow{d}
    & ? \\
    %\arrow{rr}&& 0\\
? \arrow{r}  & \Hh^1(\Weil{},\Hh^1(C,\EEx)) \arrow{d} \arrow{r}
    & \Hh^1(\Weil{},\Hh^1(B,\EEx)) \arrow{d} \arrow{r}
    & \Hh^1(\Weil{},\Hh^1(A,\EEx)) \arrow{d} \arrow{r}  &  ? \\
    &  0 & 0 & 0 & \\
  \end{tikzcd}
 \]

\subsection{Map to cohomology}\label{sec:S}
